---
title: "house_prices"
date: "`r Sys.Date()`"
author: Michael Mayer
output:
  html_document:
    toc: true
---

```{r setup, warning = FALSE, message = FALSE, echo = FALSE}
# Set initial_run to FALSE if only plots and text has to be adapted. This will save
# a lot of time as partial dependence plots are computationally demanding to calculate.

source("../r/functions.R", encoding = "UTF-8")

knitr::opts_chunk$set(fig.width = 16, fig.height = 10, echo = FALSE, 
                      warning = FALSE, message = FALSE) 

# Path to data
path <- "../rdata"

# Load prep data plus further constants
load(file.path(path, "prep.RData"))
load(file.path(path, "models.RData"))

library(tidyverse)
library(ggpubr)
library(moderndive)
library(kableExtra)

```

Blabla

## Data

We use a data set `house_prices` with information on `r format(nrow(prep), big.mark = "'")` sold houses for King County. It is shipped with R package `moderndive`.

The first six observations look as follows:

```{r}
kable(head(house_prices))
```

## Methods

## Models and their performance

We have considered the following modelling techniques

- linear regression,

- random forests and

- boosted trees.

The models were calculated on 70% of the data (training data), were validated and interpreted on 20% (validation data). 10% are still untouched for modelling (test data). The validation data set has not been used otherwise for modelling. E.g. the parameters of the boosted trees were chosen by five-fold cross-validation on the training data.

We have ended up with the following metrics evaluated on the validation data in the log world. 

```{r}
ggplot(perfTable, aes(x = label, y = value, fill = label)) + 
  geom_bar(position = position_dodge(), stat = "identity", show.legend = FALSE) +
  facet_wrap(~type) +
  ggtitle("Model performance") + 
  theme_bw(base_size = 25) +
  theme(legend.title = element_blank(), 
        legend.position = "bottom", 
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = element_blank(), y = element_blank()) +
  scale_fill_viridis_d(begin = 0.1, end = 0.7)
```



## Association of variables with response
The following diagram shows total loss increase due to random shuffling each variable. The more important the variable, the larger the loss increase.

```{r}
ggplot(varImp, aes(x = reorder(variable, dropout_loss), y = dropout_loss, fill = label)) + 
  geom_bar(position = position_dodge(), stat = "identity", show.legend = FALSE) +
  facet_wrap(~label) +
  coord_flip() +
  theme_bw(base_size = 25) +
  theme(axis.text.x = element_blank()) +
  labs(x = element_blank(), y = element_blank()) +
  scale_fill_viridis_d(begin = 0.1, end = 0.7)
```

Best predictors are `r paste(head(most_relevant), collapse = ", ")`.

## Effect per variable

We now describe the association for each covariable by considering

- **empirical effects**, i.e. average response per level of the covariable,

- corresponding partial dependence plots. They show **modelled average effects** that cannot be explained by other variables.

The results are ordered by variable importance.

```{r, results = "asis"}

for (v in most_relevant) { # v <- most_relevant[[1]]
  cat(sprintf("### %s \n", v))
  
  resp <- ggplot(mapping = aes(x = factor(temp_), y = price)) +
    stat_summary(data = eff[[v]]$data, fun.data = box_stats, geom = "crossbar",
                 width = 0.5, fill = "darkblue", colour = "black", alpha = 0.1) +
    geom_line(data = eff[[v]]$modelled, aes(group = label, color = label), size = 1.5) +
    geom_point(data = eff[[v]]$modelled, aes(group = label, color = label), size = 4) +
    theme_bw(base_size = 25) +
    theme(legend.position = c(0.5, 0.9), legend.title = element_blank(),
          legend.background = element_rect(fill = "transparent"),
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab(element_blank()) +
    scale_y_continuous(name = "Price", labels = scales::comma) +
    scale_color_viridis_d(begin = 0.1, end = 0.7)
  
  counts <- ggplot(eff[[v]]$counts, aes(x = temp_, y = n,
                   label = format(n, big.mark = "'", scientific = FALSE))) +
    geom_bar(stat = "identity", fill = "darkblue", alpha = 0.1, width = 0.5) +
    geom_text(aes(y = 0), angle = 90, hjust = -0.1, size = 6) +
    theme_void() +
    theme(strip.text.x = element_blank(), panel.grid = element_blank())

  print(ggarrange(counts, resp, heights = c(0.2, 1), ncol = 1, nrow = 2, align = "v"))
  
  cat("\n\n")
}
```
